From Nested Virtualization Architecture \\To 2D Security Management for Distributed Clouds

### 2014-03-26 paper/draft.tex
Updated comparison section.

\subsection {Limitations}

Despite the positives, MH approaches had demonstrated some limitations. Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions, inherited from its ancestors microkernels. Fortunately, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.

@@

The adoption of an extra layer of virtualization is capable to introduce a remarkable set of new features but it does not reduce the dimension of Trusted Code Base (TCB).
MH could represent a new approach to improve the potential benefits provided by the adoption of nested virtualization. In the following subsections, it will presented the four possible architectures achievable ``nesting'' of GPH and MH, underlying potentialities, drawbacks and potential cost of implementation.

@@

The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers (horizontal requirements DP1).
Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH satisfies all horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.

@@

MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach. The tiny core could be able to overcome monolithic problems due to privilege level of administration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3). However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH. Moreover, several horizontal properties are theoretically achievable but they will require a huge effort in developing functions of interest. Moreover, the lack of nested VMX functionalities prevents to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create a nested HVM guests.
To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not provide any improvement in this sense.

@@

The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.  Existing GPH-based platforms could potentially exports nested VMX features to those guests. Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the experimental status of those features.

>>>>

\noindent Despite the positives, MH approaches had demonstrated some limitations. Firstly, the straightforward one is the initial lack of hardware support, lack of device drivers are still a relevant Achilles' heel for this solutions, inherited from its ancestors microkernels. Fortunately, this kind of issue could be addressed leveraging I/O passthrough techniques, although it will cause exclusive allocation issues.

@@

\noindent Adding an extra layer of virtualization enables to introduce a wide range of new features in the IaaS infrastructure, but does not reduce the TCB size. As already discussed, introducing MH design within a nested IaaS architecture could represent a new approach to improve NV benefits by overcoming such limitations. In what follows, we analyze the four possible architectures to achieve nesting of GPH and MH, highlighting their potential, and assessing their benefits and weaknesses w.r.t. to the design principles of Section~\ref{ref:challenges}, as well as their cost of implementation.

@@

\noindent The first analyzed architecture concerns only GPH. At first, it is important to recall the most interesting achievable by such architectures. In particular, several remarkable features are live migration over different platform, transparent migration of virtualized clusters between different data-centers (horizontal requirements DP1).
Supposing the completeness and the correctness of nested virtualization hardware facilities in widespread hypervisor, GPH over GPH satisfies all horizontal design principles, but it lacks in vertical DPs as the size of TCB is the same of currently deployed architectures.

@@

\noindent MH on MH in reported only to see the orthogonality between this approach and GPH over GPH.
A theoretical implementation of MH over MH will provide the vertical requirements that are missing in the previous approach. The tiny core could be able to overcome monolithic problems due to privilege level of administration tasks (DP1), the possibility to improve the resilience of the platform to a large set of software failures (DP2) and finally reduce the possibility of the vulnerabilities of the most privilege part of the software (DP3). However, the interest on such architecture is limited because it is not able to overcome any cited issues of single-layer MH. Moreover, several horizontal properties are theoretically achievable but they will require a huge effort in developing functions of interest. Moreover, the lack of nested VMX functionalities prevents to concretely test this kind of architecture, because MH normally rely on VMX facilities in order to create a nested HVM guests.
To conclude, nested virtualization is exploited in order to fill eco-systems gaps with GPH but MH over MH does not provide any improvement in this sense.

@@

\noindent The third architecture, proposed in order to reduce security concerns, is built combining an GPH at L0 and a MH at L1. 
%Several MH are proposed in literature focusing on different .
%Therefore, this model tries to exploit microhypervisor as L1 hypervisor in order to enforce better isolation between L2 and L0. L1 will exploit the emulated nested VMX provided by GPH L0 and virtualize L2. 
%With this approach, we try to isolate the attacker in a VM virtualized on a secured layer, in order to prevent attacks to L0. The attacker would have to face with an essential virtualization layer and it would not be able to attack the back-end virtualization layer. 
An interesting point of this approach is the fact of being easily be deployed, without requiring any extra code addition to the GPH, L0.  Existing GPH-based platforms could potentially exports nested VMX features to those guests. Even if this is not so usual in modern cloud platform, this lack is considered temporary and related only to the experimental status of those features.
